Old is Still Gold: A Comparison of Cyber and Traditional Consumer Fraud in The United States

### 2017-03-13 paper/paper.tex
full intro/conclusion pass

While cybercrime is a relatively recent human invention, fraudulent activity
has a much longer history. Understanding the extent to which cybercrime is fundamentally different from, or similar to, traditional fraudulent activity is the first step toward being able to effectively combat it. This paper investigates the differences in fraud
reporting and targeting among United States based consumers by comparing
fraudulent activity mediated by the Internet with traditional forms of
fraud. By partitioning the FTC Consumer Sentinel complaint dataset into
cyber and traditional categories, we summarize the difference in fraud
trends across time, distance and location metrics. We also evaluate fraud
variations across different ethnic and socioeconomic factors. Our findings
suggest that reports of both categories of frauds decrease during
the holiday season. Individuals who are victimized via traditional methods
are more likely to complain using traditional methods, while a majority of
cyber victims submit fraud complaints online. By evaluating the top frauds
in each category, we demonstrate how traditional fraudsters have evolved in
their methodologies to execute scams on the Internet that were initially
perpetrated through traditional means.

@@

In the United States, a total of more than 25 million people are victims of frauds per year \cite{anderson2013consumer}. These deceptive scams are a major cause of economic damages to users, not to mention the added externalities of wasted time and stress. Such illicit practices have been a part of the underground economy for quite a while. Initially, individuals were tricked by scam calls, in-person fraudsters or snail mail, however, the rise of the Internet has provided fraudulent entities with a more streamlined exposure to the overall population. According to a Javelin study in 2017, identity frauds hit a record high, targeting 15.4 million victims in the United States \cite{identiytheftonline}. The findings indicated that individuals with an online presence were more susceptible to possibility of theft. Another recent study released by the  Federal Trade Commission (FTC) reported that debt collection, identity theft, and impostor scams contributed towards 56\% of the total frauds complaints in 2015 \cite{ftcpress2016} . With the number of Internet users on the rise, the number of cyber frauds is likely to increase over the next couple of years \cite{perharassment}. To deal with this rising situation, the Federal Bureau of Investigation (FBI), has also been actively involved in addressing this issue. Their complaint portal, known as the IC3 is specifically designed to record complaints on Internet-based fraudulent practices \cite{fbiic3}. This emerging trend of deceptive practices necessitates their study in order to evaluate and mitigate the harm caused to victims.

This paper evaluates the nature of consumer fraud in the United States. Our work provides a comparison between cyber and traditional frauds. We categorize cyber frauds as all those deceptive practices that victimize users online, while regular (or traditional) ones comprise of frauds that target individuals over the phone, with mail, or in-person. While we understand that online scams \cite{miramirkhani2016dial} are a major focus of today's research, our comparison based approach allows us to better understand how they differ from traditional fraud methods. It also enables us to independently evaluate trends in traditional frauds and to see whether fraudsters are adopting online mechanisms to target more individuals. This combined analysis enables us to provide strategic suggestions in means to develop better fraud reduction methodologies.

To evaluate fraud trends, we use the FTC's Consumer Sentinel database, a dataset of complaints from the year 2013 and 2014. We also collect demographic information from the US Census Bureau \cite{uscensus}. In addition to data collection, we devise a calibration methodology to identify and separate cyber frauds from regular ones in the complaint dataset.

@@

The goal of this paper is to investigate the dynamics of cyber and traditional methods for committing fraud in the United States. By partitioning the FTC Consumer Sentinel complaint dataset, we are able to explore trends among cyber and regular fraud, and analyze trends along time, distance and location metrics as well as their variation across consumer demographics. While the Internet has greatly expanded the potential reach and scale of fraudsters: where once they could only use mail or telephones for various frauds, the Internet enables one to contact millions of users very quickly, whether through fraudulent websites, Craigslist posts, or spam emails. Even so, the difference between the cyber and non-cyber methodologies is not overwhelming: this effect suggests that non-cyber activities, like purchasing goods or posting money orders, may still serve as a limiting factor on the extent or targeting of these fraudulent activities.

Fraud analysis is an impactful topic that has been understudied in academia. While our current work has its limitations, we plan to extend our analysis over multiple years by establishing a collaboration with the FTC and obtaining high fidelity datasets that span a longer duration. We also look forward to evaluating our current findings with data collected by other regulatory agencies such as the FBI. This will eventually also allow us to overcome certain nuances and limitations of the current evaluation and help us develop stronger insights. Another interesting future direction is to evaluate how international fraud compares to fraud within the US, and identify any prominent international fraudsters who target US victims. This can possibly be achieved by using data from the European Commission and the eCrime project.

@@

We would like to thank the reviewers for reviewing our work and providing
helpful feedback. This material is based upon work supported by the National Science Foundation
 under Grant No. 1351058. Any opinions,
findings, and conclusions or recommendations expressed in this material are
those of the authors and do not necessarily reflect the views of the
National Science Foundation.

>>>>

While cybercrime is a relatively recent human invention, fraudulent activity
has a much longer history. Understanding the extent to which cybercrime is
fundamentally different from, or similar to, traditional fraudulent activity is
the first step toward being able to effectively combat it. This paper
investigates the differences in fraud targeting among United States based
consumers by comparing reports of fraudulent activity mediated by the Internet
with traditional forms of fraud. By partitioning the FTC Consumer Sentinel
complaint dataset into cyber and traditional categories, we summarize the
difference in fraud trends across time, distance and location metrics. We also
evaluate fraud variations across different ethnic and socioeconomic factors.
Our findings suggest that reports of both categories of frauds decrease during
the holiday season. Individuals who are victimized via traditional methods are
more likely to complain using traditional methods, while a majority of cyber
victims submit fraud complaints online. By evaluating the top frauds in each
category, we demonstrate how traditional fraudsters have evolved in their
methodologies to execute scams on the Internet that were initially perpetrated
through traditional means.

@@

In the United States, a total of more than 25 million people are victims of
frauds per year \cite{anderson2013consumer}. These deceptive scams are a major
cause of economic damages to users, as well as added stress and wasted time.
Such illicit practices have been a part of the underground economy for quite a
while. Initially, individuals were contacted via scam calls,  snail mail, or in
person communication.  The rise of the Internet has provided fraudulent
entities with a more effective method of communicating with the overall
population. According to a Javelin study in 2017, identity frauds hit a record
high, targeting 15.4 million victims in the United
States~\cite{identiytheftonline}. The findings indicated that individuals with
an online presence were more susceptible to possibility of theft. Another
recent study released by the  Federal Trade Commission (FTC) reported that debt
collection, identity theft, and impostor scams contributed towards 56\% of the
total frauds complaints in 2015 \cite{ftcpress2016} . With the number of
Internet users on the rise, the number of cyber frauds is likely to increase
over the next couple of years \cite{perharassment}. To deal with this rising
situation, the Federal Bureau of Investigation (FBI), has also been actively
involved in addressing this issue. Their complaint portal, known as the IC3 is
specifically designed to record complaints on Internet-based fraudulent
practices \cite{fbiic3}. This emerging trend of deceptive practices
necessitates their study in order to evaluate and mitigate the harm caused to
victims.

This paper evaluates the nature of consumer fraud in the United States. Our
work provides a comparison between cyber and traditional frauds. We categorize
cyber frauds as all those deceptive practices that victimize users online,
while regular (or traditional) frauds target individuals over the phone, via
mail, or in-person. While we understand that online
scams~\cite{miramirkhani2016dial} are a major focus of today's research, our
comparison based approach allows us to better understand how they differ from
traditional fraud methods. It also enables us to independently evaluate trends
in traditional frauds and to determine whether fraudsters are adopting online
mechanisms to target more individuals. This combined analysis enables us to
provide strategic suggestions in means to develop better fraud prevention
techniques.

To evaluate fraud trends, we use the FTC's Consumer Sentinel database, a dataset of complaints from the year 2013 and 2014. We also collect demographic information from the US Census Bureau~\cite{uscensus}. In addition to data collection, we devise a calibration methodology to identify and separate cyber frauds from regular ones in the complaint dataset.

@@

The goal of this paper is to investigate the dynamics of cyber and traditional
methods for committing fraud in the United States. By partitioning the FTC
Consumer Sentinel complaint dataset, we are able to explore trends among cyber
and regular fraud, and analyze trends along time, distance and location metrics
as well as their variation across consumer demographics. The Internet has
greatly expanded the potential reach and scale of fraudsters, enabling them
to contact millions of users very quickly, whether through fraudulent websites,
Craigslist posts, or spam emails. Even so, the difference between the cyber and
non-cyber methodologies is not overwhelming: this effect suggests that
non-cyber activities, like purchasing goods or posting money orders, may still
serve as a limiting factor on the extent or targeting of these fraudulent
activities.

As a future effort in this space, we plan to extend our analysis over multiple
years through analysis of different, higher fidelity datasets that span a
longer duration. We also look forward to evaluating our current findings with
data collected by other regulatory agencies such as the FBI. This will
eventually also allow us to overcome certain nuances and limitations of the
current evaluation and help us develop stronger insights. Another interesting
future direction is to evaluate how international fraud compares to fraud
within the US, and identify any prominent international fraudsters who target
US victims. This can possibly be achieved by using data from the European
Commission and the eCrime project.

@@

We would like to thank the reviewers for their helpful feedback. This material
is based upon work supported by the National Science Foundation under Grant No.
1351058. Any opinions, findings, and conclusions or recommendations expressed
in this material are those of the authors and do not necessarily reflect the
views of the National Science Foundation.
